---
title: "Fedorov Experiments"
output: html_notebook
---
```{r warning=FALSE}
# Load all components
here <- dirname(rstudioapi::getSourceEditorContext()$path)
source(paste(here,"FedorovDesignClass.R",sep="/")) # ignore error
source(paste(here,"Fedorov.R",sep="/")) 
source(paste(here,"FedorovGA.R",sep="/")) 
library(AlgDesign)
library(tidyverse)
```

# Controls
``` {R}
# experiment size
size <- 'small'

# constraints
lmda <- 0
iter <- 100
```


# Initialize DesignMatrix for small experiment

```{r}
# how many cards?
if (size == 'small'){
  n <- 8
} else if (size == 'medium') {
  n <- 16
} else if (size == 'large') {
  n <- 40
} else {
  stop('Error in "size"')
}

# create DesignMatrix object
dm <- DesignMatrix(n=n, cholesky=T)

# create attributes
## base-case for 'small' design
dm$add_attribute(name="age", levels=3, dist=c(25,50,25))
dm$add_attribute(name="gender", levels=2, dist=c(50,50))
dm$add_attribute(name="bmi", levels=3, dist=c(25,25,50))

## add add'l attributes for med/large
if (size == 'medium') {
  dm$add_attribute(name="race", levels=4, dist=c(40,20,20,20))
  dm$add_attribute(name="dia", levels=2, dist=c(50,50))
  dm$add_attribute(name="ldl", levels=3, dist=c(25,50,25))
  dm$add_attribute(name="bp", levels=3, dist=c(25,50,25))
  dm$add_attribute(name="a1c", levels=4, dist=c(25,25,25,25))
}
if (size == 'large') {
  dm$add_attribute(name="angn", levels=3, dist=c(50,25,25))
  dm$add_attribute(name="ren", levels=3, dist=c(50,25,25))
  dm$add_attribute(name="srcl", levels=3, dist=c(50,25,25))
  dm$add_attribute(name="uacr", levels=3, dist=c(33,33,34))
  dm$add_attribute(name="strk", levels=2, dist=c(75,25))
  dm$add_attribute(name="ptx", levels=4, dist=c(25,25,25,25))
  dm$add_attribute(name="hist", levels=3, dist=c(50,25,25))
  dm$add_attribute(name="smk", levels=3, dist=c(50,25,25))
}

# add interactions
## no interactions for 'small' design
if (size == 'medium') {
  dm$add_interaction(n1="dia", n2="a1c", l1=0, l2=3, eq=F) # nondiabetics can't have high a1c
  dm$add_interaction(n1="dia", n2="a1c", l1=1, l2=0, eq=F) # diabetics can't have low a1c
}
if (size == 'large') {
  dm$add_interaction(n1="bp", n2="ptx", l1=3, l2=0, eq=F) # patients with high bp cannot be untreated
}

### generate candidate set 
candidate_set <- gen.factorial(dm$levels, factors="all")
# convert to numeric matrix
indx <- sapply(candidate_set, is.factor)
candidate_set[,indx] <- lapply(candidate_set[,indx], function(x) as.numeric(as.character(x)))
candidate_set <- as.matrix(candidate_set)
# zero base
candidate_set <- candidate_set-1

### generate design
dm$generate_design()

cat("Example Design Class Object\n--------\n")
cat("names: \n", dm$names, "\n")
cat("\nvalues: \n")
dm$values
cat("\ninteraction slacks: \n")
dm$islacks
cat("\ndistribution slacks: \n")
dm$dslacks
cat("\ndesign matrix: \n")
dm$X
```

# Long run test

```{r warning=FALSE, include=FALSE}
out_df <- data.frame()

# loop conditions
trial <- 1
attempts <- 1000
while (trial <= attempts) {
  print(paste("Trial: ", trial))
  dm$generate_design()
  
  # duplicate design matrix for testing
  dm_fed <- dm$copy(shallow=T)
  dm_fed$set_cholesky(cholesky=F)
  dm_ga <- dm$copy(shallow=T)
  
  # test initial optimality
  init_dopt <- doptimality(dm, lambda=lmda, how='det')
  
  # run algos
  f_time <- system.time({
    fed_out <- tryCatch(
      expr = {fedorov(dm_fed, candidate_set, n, lambda=lmda, iter=iter, return_iter=TRUE)},
      error = function(e) {
        message("Error in Fedorov: ")
        print(dm$X)        
        message(paste("Returns error: ", e))
        return(NULL)
      }
    )
    if (is.null(fed_out)) {
      message(paste("Restarting loop: ", trial))
      next
    } 
  })
  f_DM <- fed_out[[1]]
  f_iters <- fed_out[[2]]  
  f_dopt <- doptimality(f_DM, lambda=lmda, how='det')

  ga_time <- system.time({
    ga_out <- tryCatch(
      expr = {gen_alg(dm_ga, pop=16, gens=1000, test='doptimality', lambda=lmda, return_iter=TRUE)},
      error = function(e) {
        message("Error in GA: ")
        print(dm$X)
        message(paste("Returns error: ", e))
        return(NULL)
      }
    )
    if (is.null(fed_out)) {
      message(paste("Restarting loop: ", trial))
      next
    }
  })
  
  ga_DM <- ga_out[[1]]
  ga_iters <- ga_out[[2]]      
  ga_dopt <- doptimality(ga_DM, lambda=lmda, how='det')

  # preserve run data
  fedorov_out <- c(trial, 'fedorov', init_dopt, f_iters, f_time[['elapsed']], f_dopt)
  out_df <- rbind(out_df, fedorov_out, stringsAsFactors=FALSE)
  genetic_out <- c(trial, 'genetic', init_dopt, ga_iters, ga_time[['elapsed']], ga_dopt)
  out_df <- rbind(out_df, genetic_out, stringsAsFactors=FALSE)
  
  trial <- trial + 1
}

colnames(out_df) <- c('trial', 'algo', 'init_dopt', 'iters', 'time', 'final_dopt')
out_df$init_dopt <- as.numeric(out_df$init_dopt)
out_df$iters <- as.numeric(out_df$iters)
out_df$time <- as.numeric(out_df$time)
out_df$final_dopt <- as.numeric(out_df$final_dopt)
write.csv(out_df, paste0('/Users/mithras/Desktop/FedorovTest_lmda', lmda, '_', size, '.csv'), row.names=FALSE)

print(paste('Test with lmda =', lmda, 'completed. CSV saved.'))
```

# Stats
``` {R}
size <- 'small'
```

## Run stats on tests with lmda = 1
* average # iterations
* average time to solve
* averate time / iteration
* average d-optimality of sol'n

```{r}
if (!exists('lmda1_df')) {
  lmda1_df <- read.csv(paste0('/Users/mithras/Desktop/FedorovTest_lmda1_',size,'.csv'), header=TRUE)
}
lmda1_df

lmda1_df$itertime <- lmda1_df$time / lmda1_df$iters

lmda1_df %>%
  group_by(algo) %>%
  summarize(
    iters = mean(iters),
    time = mean(time),
    itertime = mean(itertime),
    final_dopt = mean(final_dopt)
  )

```

## How many times did fedorov outperform GA?
``` {R}
lmda1_df %>%
  group_by(algo) %>%
  count(final_dopt)

```


## Sig. testing for differences btwn algos

```{R}
print('T-test for equivalence of average runtime,  If p < 0.5, reject.')
t.test(time ~ algo, data = lmda1_df, paired=TRUE)$p.value

print('T-test for equivalence of average final doptimality,  If p < 0.5, reject.')
t.test(final_dopt ~ algo, data = lmda1_df, paired=TRUE)$p.value
```


## Run stats on tests with lmda = 0
* average # iterations
* average time to solve
* averate time / iteration
* average d-optimality of sol'n

```{r}
if (!exists('lmda0_df')) {
  lmda0_df <- read.csv(paste0('/Users/mithras/Desktop/FedorovTest_lmda1_',size,'.csv'), header=TRUE)
}
lmda0_df

lmda0_df$itertime <- lmda0_df$time / lmda0_df$iters

lmda0_df %>%
  group_by(algo) %>%
  summarize(
    iters = mean(iters),
    time = mean(time),
    itertime = mean(itertime),
    final_dopt = mean(final_dopt)
  )

```

## How many times did fedorov outperform GA?
``` {R}
lmda0_df %>%
  group_by(algo) %>%
  count(final_dopt)

```


## Sig. testing for differences btwn algos

```{R}
print('T-test for equivalence of average runtime,  If p < 0.5, reject.')
t.test(time ~ algo, data = lmda0_df, paired=TRUE)$p.value

print('T-test for equivalence of average final doptimality,  If p < 0.5, reject.')
t.test(final_dopt ~ algo, data = lmda0_df, paired=TRUE)$p.value

```


```{r}
```


```{r}
```


```{r}
```

